### 2014-09-22 Studies/2011_San_Francisco/Report/San_Francisco_Fire.tex
San Francisco: Minor edits to report

Input data from various sources must be collected and documented to simulate a fire using FDS or any other fire model. For the simulation results presented in this document, information was obtained from two primary sources. The following information was gathered from the fire scene: the geometry of the building and the compartments being modeled, the size and location of exterior and interior ventilation openings, and documentation of fire damage to the building. The following information was gathered from witnesses, first responders, reports, and recorded media such as fire ground radios or videos: information on the timing of the fire development, the sequence and approximate timing of ventilation openings to the outside, and weather conditions at the time of the fire.

@@

To estimate the type of fuel, fire size, and structure ventilation, on-scene videos and post-incident reports were used. Based on these sources of information, the FDS model includes three source fires: an initial couch fire in the basement, a secondary flashover fire that involved additional furniture items in the basement, and a third fire that occurred on the exterior basement balcony. The basement primarily contained upholstered polyurethane furniture fuel items, and the balcony was constructed of wood. The HRR of the initial couch fire is based on an experimentally measured HRR for a mattress fire~\cite{madrzykowski2009fire} to represent the fire growth rate of a similar household furniture item. The primary fuel packages were located in the basement as shown in Fig.~\ref{fig:fuel_placement}. In this figure, the numbered labels indicate the order in which the fuel items were burning: 1) the initial couch fire, 2) the secondary furniture fuel items, and 3) the balcony fire.

@@

Figure~\ref{fig:hrr} shows the prescribed design fire for this scenario, which is the sum of the HRRs of the three source fires. In this figure, the vertical lines indicate the three stages of the fire based on the contribution of the fuel packages. Therefore, the estimated maximum specified fire size for this structure was approximately 20~MW. Note that the prescribed HRR can be different than the simulated HRR calculated by FDS based on the amount of available fuel, oxygen, and ventilation. A comparison of the prescribed HRR and the simulated HRR is discussed in Section~\ref{sec:HRR}.

\begin{figure}[!ht]
\includegraphics[width=5.5in]{../Figures/Fire_HRR}
\caption[Prescribed HRR vs. time.]
{Prescribed HRR vs. time. The vertical lines indicate the three stages of the fire based on the contribution of the fuel packages.}
\label{fig:hrr}
\end{figure}

@@

The simulations in this study account for changes in the ventilation due to a combination of fire department operations (opening doors) and fire acting on the structure (fire breaching walls, breaking windows, etc). The ventilation times and ventilation areas represent our best understanding of the incident. The times of the ventilation changes of exterior doors and windows are provided in Table~\ref{tab:ventilation_timeline}.

During the time at which the downed firefighters were located, the state of the windows and doors in the structure was known, which resulted in the establishment of a flow path in the interior stairwell leading into the basement. However, the exact time of operation for some doors was not known. Therefore, these doors were assumed to be open at the start of the simulation, which included 1) the front door, 2) the interior door leading from the hallway on the first floor into the stairwell towards the basement, 3) the door leading from the first floor landing into the garage, and 4) the overhead garage door.

>>>>

Input data from various sources must be collected and documented to simulate a fire using FDS or any other fire model. For the simulation results presented in this document, information was obtained from two primary sources. The following information was gathered from the fire scene: the geometry of the building and the compartments being modeled, the size and location of exterior and interior ventilation openings, and documentation of fire damage to the building. The following information was gathered from witnesses, first responders, reports, and recorded media such as fire ground radios or photographs: information on the timing of the fire development, the sequence and approximate timing of ventilation openings to the outside, and weather conditions at the time of the fire.

@@

To estimate the type of fuel, fire size, and structure ventilation, on-scene photographs and post-incident reports were used. Based on these sources of information, the FDS model includes three source fires: an initial couch fire in the basement, a secondary flashover fire that involved additional furniture items in the basement, and a third fire that occurred on the exterior basement balcony. The basement primarily contained upholstered polyurethane furniture fuel items, and the balcony was constructed of wood. The HRR of the initial couch fire is based on an experimentally measured HRR for a mattress fire~\cite{madrzykowski2009fire} to represent the fire growth rate of a similar household furniture item. The primary fuel packages were located in the basement as shown in Fig.~\ref{fig:fuel_placement}. In this figure, the numbered labels indicate the order in which the fuel items were burning: 1) the initial couch fire, 2) the secondary furniture fuel items, and 3) the balcony fire.

@@

Figure~\ref{fig:hrr} shows the prescribed design fire for this scenario, which is the sum of the HRRs of the three source fires. In this figure, the text labels indicate the three stages of the fire based on the contribution of the fuel packages. Therefore, the estimated maximum specified fire size for this structure was approximately 20~MW. Note that the prescribed HRR can be different than the simulated HRR calculated by FDS based on the amount of available fuel, oxygen, and ventilation. A comparison of the prescribed HRR and the simulated HRR is discussed in Section~\ref{sec:HRR}.

\begin{figure}[!ht]
\includegraphics[width=5.5in]{../Figures/Fire_HRR}
\caption[Prescribed HRR vs. time.]
{Prescribed HRR vs. time. The text labels indicate the three stages of the fire based on the contribution of the fuel packages.}
\label{fig:hrr}
\end{figure}

@@

The simulations in this study account for changes in the ventilation due to a combination of fire department operations (opening doors) and fire acting on the structure (fire breaching walls, breaking windows, etc). The ventilation times and ventilation areas represent our best understanding of the incident. The times of ventilation changes for doors and windows are provided in Table~\ref{tab:ventilation_timeline}.

During the time at which the downed firefighters were located, the state of some windows and doors in the structure was known, which resulted in the establishment of a flow path in the interior stairwell leading into the basement. However, the exact time of operation for other doors was not known. Therefore, the following doors were assumed to be open at the start of the simulation: 1) the front door, 2) the interior door leading from the hallway on the first floor into the stairwell towards the basement, 3) the door leading from the garage into the first floor landing, and 4) the overhead garage door. The basement windows and exterior door were specified to open at certain times based on observations of failure (i.e., window breakage) or fire department operations (i.e., forcible entry) from on-scene photographs and post-incident reports.
